---
Name: Mason Shu
Date: 8/18/2024
STA 380, Part 2: Exercises
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries


```

# Probability practice
## Part A
```{r, echo=FALSE}
# The fraction of Random Clickers (RC) is P(RC) = 0.3, so the fraction of Truthful Clickers (RC) is P(TC) = 1 - P(RC) = 0.7
# The probability that a Random Clicker answers "Yes" is P(Yes|RC) = 0.5, and the probability that they answer "No" is P(No|RC) = 0.5
# The overall fraction of people who answered "Yes" is 65%, so P(Yes) = 0.65
# Find probability that a Truthful Clicker answered "Yes": P(Yes|TC)

# Total Probability: P(Yes) = P(Yes|RC) * P(RC) + P(Yes|TC) * P(TC)
# 0.65 = 0.5 * 0.3 + P(Yes|TC) * 0.7
(0.65 - (0.5 * 0.3))/0.7

```

**Given from the problem:**

- The fraction of Random Clickers (RC) is P(RC) = 0.3, so the fraction of Truthful Clickers (RC) is P(TC) = 1 - P(RC) = 0.7

- The probability that a Random Clicker answers "Yes" is P(Yes|RC) = 0.5, and the probability that they answer "No" is P(No|RC) = 0.5

- The overall fraction of people who answered "Yes" is 65%, so P(Yes) = 0.65

**Find the probability that a Truthful Clicker answered "Yes": P(Yes|TC)**

- Total Probability: P(Yes) = P(Yes|RC) * P(RC) + P(Yes|TC) * P(TC)

- 0.65 = 0.5 * 0.3 + P(Yes|TC) * 0.7

- P(Yes|TC) = 0.7142857

The fraction of people who are truthful clickers that answered yes is **0.7142857**, or approximately 71.43%.

## Part B
```{r, echo=FALSE}
# P(Positive|Disease) = 0.993
# P(Negative|No Disease) = 0.9999, so P(Positive|No Disease) = 1 - 0.9999 = 0.0001
# P(Disease) = 0.000025, so P(No Disease) = 1 - 0.000025 = 0.999975
# Find probability that someone has the disease given that they tested positive: P(Disease|Positive)

# Bayes Theorem: P(Disease|Positive) = (P(Positive|Disease) * P(Disease)) / P(Positive)
# Total Probability: P(Positive) = P(Positive|Disease) * P(Disease) + P(Positive|No Disease) * P(No Disease)

# P(Positive) = (0.993 * 0.000025) + (0.0001 * 0.999975)
ppositive = (0.993 * 0.000025) + (0.0001 * 0.999975)

# P(Disease|Positive) = (0.993 * 0.000025) / ppositive
(0.993 * 0.000025) / ppositive

```

**Given from the problem:**

- P(Positive|Disease) = 0.993

- P(Negative|No Disease) = 0.9999, so P(Positive|No Disease) = 1 - 0.9999 = 0.0001

- P(Disease) = 0.000025, so P(No Disease) = 1 - 0.000025 = 0.999975

**Find probability that someone has the disease given that they tested positive: P(Disease|Positive)**

- Bayes Theorem: P(Disease|Positive) = (P(Positive|Disease) * P(Disease)) / P(Positive)

- Total Probability: P(Positive) = P(Positive|Disease) * P(Disease) + P(Positive|No Disease) * P(No Disease)

- P(Positive) = (0.993 * 0.000025) + (0.0001 * 0.999975)

- ppositive = (0.993 * 0.000025) + (0.0001 * 0.999975) = 0.0001248225

- P(Disease|Positive) = (0.993 * 0.000025) / ppositive

- (0.993 * 0.000025) / ppositive = 0.1988824

The probability that someone has the disease given that they test positive is **0.1988824**, or approximately 19.89%.

